---
title: "PA GMB"
author: "Christopher Huong"
date: "2023-12-10"
output: html_document
---

Analysis code for 

"Associations between physical activity and subcategories of mental health: A propensity score analysis among a global sample of 341,956 adults"

Christopher Huong & Denver Brown 2023

Simplified and annotated for resubmission 12/10/2023

```{r, warning=F, message=F}
library(tidyverse)
library(naniar)
library(WeightIt)
library(survey)
library(cobalt)

```

Load in raw data, check variable names, and check range of response dates

```{r}
load('rawdata.RData')

names(rawdata)
rawdata$Submit.Date..UTC. <- as.Date(rawdata$Submit.Date..UTC.)
range(rawdata$Submit.Date..UTC.)

```

Select DVs, predictor, and covariates, rename variables
```{r}
cleandata <- rawdata %>% 
  select(Overall.MHQ, Cognition, Adaptability...Resilence, Drive...Motivation, Mood...Outlook, Social...Self, Mind.Body.Connection, Frequency.of.doing.exercise, Age, Biological.Sex, Gender.Identity, Country, Ethnicity, Education, Employment, Current.Family.Situation, Frequency.of.Socializing, Frequency.of.getting.a.good.nights.sleep, Presence.Absence.of.Diagnosed.Medical.Disorder, Mental.Health.Treatment.Status, Childhood.traumas, Adult.traumas)

names(cleandata) <- c('mhq', 'cog', 'adaptresil', 'drivemotiv', 'moodoutlook', 'socialself', 'mindbody', 'PA', 'age', 'sex', 'genderidentity', 'country', 'ethnicity', 'education', 'employment', 'relationship', 'socialize', 'sleep', 'meddiagnosis', 'mhseeking', 'childtrauma', 'adulttrauma')
```

Preprocessing variables

```{r}
summary(cleandata$PA)
# remove 2 non-english responses
cleandata <- cleandata %>% subset(PA == "Every day" | PA == "Few days a week" | PA == "Less than once a week" | PA == "Once a week" | PA == "Rarely/Never")
# recode to binary treatment
cleandata <- cleandata %>% mutate(PA = if_else(PA == "Rarely/Never", 0L, 1L))
table(cleandata$PA)


summary(cleandata$age)
cleandata$age <- as.numeric(cleandata$age)

summary(cleandata$sex)
cleandata$sex <- factor(cleandata$sex, order=F)

summary(cleandata$genderidentity)
cleandata$genderidentity <- factor(cleandata$genderidentity, order=F)

cleandata$country <- as.integer(cleandata$country)
summary(cleandata$country)

cleandata <- cleandata %>%              
  mutate(education = case_when(education == "Primary Education" ~ "less.hs",
                               education == "Some High School" ~ "less.hs",
                               education == "Médio completo" ~ "less.hs",
                               education == "High School" ~ "hs",
                               education == "Ensino técnico" ~ "vocational",
                               education == "Ensino profissionalizante" ~ "vocational",
                               education == "Vocational certification" ~ "vocational",
                               education == "Associate’s Degree" ~ "assoc.deg",
                               education == "Bachelor's Degree" ~ "bach.deg",
                               education == "Master's Degree" ~ "grad.deg",
                               education == "PhD" ~ "grad.deg",
                               education == "J.D" ~ "grad.deg",
                               education == "J.D. (Direito)" ~ "grad.deg",
                               education == "M.D." ~ "grad.deg",
                               education == "M.D. (Medicina)" ~ "grad.deg",
                               education == "Other" ~ "other",
                               education == "Prefer not to say" ~ "Prefer not to say"))
cleandata$education <- factor(cleandata$education, order=F)
summary(cleandata$education)

cleandata$employment <- factor(cleandata$employment, order = F)
summary(cleandata$employment)

cleandata$relationship <- factor(cleandata$relationship, order = F)
summary(cleandata$relationship)

cleandata$socialize <- factor(cleandata$socialize, order = F, levels = c("Rarely/Never", "1-3 times a month", "Once a week", "Several days a week"))
summary(cleandata$socialize)

cleandata$sleep <- factor(cleandata$sleep, order = F, levels = c("Hardly ever", "Some of the time", "Most of the time", "All of the time"))
summary(cleandata$sleep)

cleandata$meddiagnosis <- factor(cleandata$meddiagnosis, order = F)
summary(cleandata$meddiagnosis)

cleandata$mhseeking <- factor(cleandata$mhseeking, order = F)
summary(cleandata$mhseeking)


cleandata <- cleandata %>%
  mutate(childtrauma = if_else(childtrauma =="|I did not experience any of the above during my childhood" | childtrauma == "|None of the above", 0L,1L),
         adulttrauma = if_else(adulttrauma == "|I did not experience any of the above" | adulttrauma == "|None of the above", 0L,1L))

table(cleandata$childtrauma)
table(cleandata$adulttrauma)


```


Code and check missingness

```{r}

colSums(is.na(cleandata)) #check NAs per var
apply(cleandata == "", 2, sum)
apply(cleandata == "Prefer not to say", 2, sum)
cleandata[cleandata == ""] <- NA
cleandata[cleandata == "Prefer not to say"] <- NA 

colSums(is.na(cleandata)) #check NAs per var

#percent missingness by var
percentmiss <- function(x){
  sum(is.na(x)) / length(x) * 100} 

apply(cleandata, 2, percentmiss)

```

"Data inspection revealed considerable missingness for ethnicity (84.2%) and gender identity (98.5%) due to only having been included on surveys for individuals who reported residing in certain countries and therefore these variables were excluded"

```{r}
cleandata <- cleandata %>% select(-c(ethnicity, genderidentity))
cleandata <- droplevels(cleandata)  #get rid of "Prefer not to say" factor levels
summary(cleandata)

# save(cleandata, file="cleandata.RData")
```


# Propensity score weighting using GBM

Check pre-weighting covariate balance across "treatment" groups
```{r}
bal.tab(PA ~ age + sex + education + employment + relationship + socialize + sleep + meddiagnosis + mhseeking + childtrauma + adulttrauma,
        data = cleandata,
        s.d.denom = "pooled",
        stats = c("m", "ks"),
        thresholds = c(m = .01))
```


Propensity score weighting using generalized boosted modeling and binary treatment based on the ATC estimand. 
```{r}

gbm_weights <- weightit(PA ~
                        age
                      + sex
                      + education
                      + employment
                      + relationship
                      + socialize
                      + sleep
                      + meddiagnosis
                      + mhseeking
                      + childtrauma
                      + adulttrauma,
                      data = cleandata,
                      method = "gbm",
                      estimand = "ATC",
                      trim.at = 0.99,
                      distribution = "bernoulli")

save(gbm_weights, file = "gbm_weights.RData")

load("gbm_weights.RData")

```

Covariate balancing diagnostics
```{r}
gbm_weights
summary(gbm_weights)


bal.tab(gbm_weights$treat ~ gbm_weights$covs,
        data = gbm_weights,
        weights = gbm_weights$weights,
        stats = c("m"),
        s.d.denom = "control",
        thresholds = c(m = .01))

love.plot(gbm_weights, binary = "std", var.order = "un", stats = "m",
           thresholds = .01) + theme(legend.position = "top")
```

Estimating main treatment effect using weighted GLM, clustered by country
```{r}
des <- svydesign(ids = ~country, weights = gbm_weights$weights, data = cleandata)



mhq <- svyglm(mhq ~ PA,
              design = des,
              family = gaussian())

summary(mhq)
# Coefficients:
#   Estimate Std. Error t value Pr(>|t|)    
#   PA            17.857      1.413   12.64   <2e-16 ***
confint(mhq)
#                  2.5 %   97.5 %
#   (Intercept) 41.86981 52.25683
# PA            15.07241 20.64148

```






















